---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: PA1_Sep2015.html
    keep_md: true
---


## Loading and preprocessing the data
```{r eval=FALSE}
library(data.table)
library(shiny)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lattice)
df <- read.csv("activity.csv")
```
## What is mean total number of steps taken per day?
Use aggregate() to make a new datatable with date v.s. total steps
```{r eval=FALSE}
total_steps <- aggregate(steps ~ date, FUN = sum, data = df)
```

The histogram of total steps per day

```{r echo=FALSE}
  hist(total_steps$steps)
```

There mean and median of total number of steps per day:
```{r echo=TRUE}
mean(total_steps$steps)
median(total_steps$steps)
```
## What is the average daily activity pattern?
Use the summarize() function to create a new datatable interval v.s. steps:
```{r eval=FALSE}
tmp <- summarize(group_by(df, interval), mean(steps, na.rm = T))
```
Average daily activity

```{r echo=FALSE}
  plot(tmp, type = "l", 
       main = "Average across all Days",
       xlab = "5-minute Interval",
       ylab = "Average Number of Steps Taken")
```

Using summary(), find the maximum number of steps. Use the result to find the row containing maximum number of steps.
```{r echo=TRUE}
summary(tmp)
filter(tmp, tmp$`mean(steps, na.rm = T)` > 206)
```

## Imputing missing values
Use summary() to find out how many rows contains missing values(NA)
```{r echo=TRUE}
summary(df)
```
--> missing value = 2304 rows

Replace NA with mean for that 5-minute interval, write the result to a tidy file called padded.csv

```{r eval=FALSE}
new <- replace_na(df, replace = list(steps = tmp$'mean(steps, na.rm = T)'))
write.csv(new, "padded.csv", row.names = F)
```

Histogram of total number of steps take each day

```{r echo=FALSE}
  tmp <- aggregate(steps ~ date, FUN = sum, data = new)
  hist(tmp$steps)
```

```{r echo=TRUE}
mean(tmp$steps)
median(tmp$steps)
```
## Are there differences in activity patterns between weekdays and weekends?

Use wday() and bind_cols() to create a new column weekdays
```{r eval=FALSE}
library(lubridate)
x <- sub('2', 'weekday',
        sub('3', 'weekday',
            sub('4', 'weekday',
                sub('5', 'weekday',
                    sub('6', 'weekday',
                        sub('7', 'weekend', 
                            sub('1', 'weekend',
                                wday(new$date))))))))
weekdays <- bind_cols(new, data.table(weekdays = x))
```

```{r echo=TRUE}
head(weekdays)
```

Calculate average steps based on weekday(group1) or weekend(group2) first and combine them back together.

```{r eval=FALSE}
group1 <- filter(weekdays, weekdays == "weekday")
b1 <- aggregate(steps ~ interval, FUN = mean, data = group1) %>%
  mutate(weekdays = "weekday")
group2 <- filter(weekdays, weekdays == "weekend")
b2 <- aggregate(steps ~ interval, FUN = mean, data = group2) %>%
  mutate(weekdays = "weekend")
AvgByTypeOfDays <- bind_rows(b1,b2)
```

Comparing weekdays and weekend

```{r echo=FALSE}
  library(lattice)
  xyplot(steps ~ interval|weekdays, 
         AvgByTypeOfDays, type = "l", layout = c(1,2))
```